Computing Exact Restricted Isometry Constants via Mixed Integer Semidefinite Programming

### 2016-03-20 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
paper: make "contribution of this work" a section
paper: minor changes

\subsection{Contribution of this Work}

@@

The idea is that the optimal constant $\alpha_k^2$ in \eqref{lowerRIP} can be computed
by the non-convex quadratic optimization problem
\begin{equation}\label{QP}\tag{QP}
 \begin{aligned}
  \min \quad & \Norm{Ax}{2}^2  \\
  \text{s.t.} \quad & \Norm{x}{2}^2 = 1,  \\
  & \Norm{x}{0} \leq k. 
 \end{aligned}
\end{equation}
The corresponding maximization problem allows to compute $\beta_k^2$. D'Aspremont et al. \cite{Asp07} tackled the non-convex quadratic equality constraint 
$\Norm{x}{2}^2 = 1$ by the technique of semidefinite lifting, first introduced by Goemans and Williamson~\cite{GW95} 
for the max-cut problem. The idea is to use new matrix variables $X=xx\T$ for the quadratic terms, where the condition $X=xx\T$ is enforced by the equivalent constraints $X \succeq 0$ and $\Rk(X) = 1$, where, as usual, 
$X \succeq 0$ means that $X$ is symmetric (shortly written $X \in S_n$) and positive semidefinite. The objective and the normalization-constraint can then be expressed as  $\Norm{Ax}{2}^2 = x\T A\T Ax =
\Tr(A\T A X)$ and $1 = \Norm{x}{2}^2 = \sum_{i=1}^n x_i^2 = \Tr(X)$.

In the next step, the inequality $\Norm{x}{0} \leq k$ is then substituted by the weaker constraint $\ones\T
\abs{X} \ones \leq k$, which follows from the fact that $\ones\T\abs{X} \ones = \Norm{\vect(X)}{1}\leq \sqrt{\Norm{\vect(X)}{0}}\, \Norm{\vect(X)}{2}$ and $\Norm{\vect(X)}{0}=\Norm{x}{0}^2$ in the rank-1 case, where $\vect(X)$ is a vector
in $\R^{n^2}$ consisting of all entries of $X$. 
Relaxing the non-convex rank constraint finally leads to the following SDP:
\begin{equation}\label{Asp07}\tag{A1}
 \begin{aligned}
  \min \quad & \Tr(A\T A X) \\
  \text{s.t.} \quad & \Tr(X) = 1, \\
  & \ones\T \abs{X} \ones \leq k, \\
  & X \succeq 0. \nonumber
 \end{aligned}
\end{equation}

Another SDP relaxation was proposed by d'Aspremont et al. in \cite{Asp08}, where
\begin{equation}\label{phi}
  \phi(\rho) = \max \big\{ x\T \Sigma\, x - \rho \Norm{x}{0} \suchthat \Norm{x}{2} \leq 1 \big\},
\end{equation}
originally stemming from sparse principal component analysis, is approximately solved for a symmetric positive semidefinite matrix $\Sigma$. For
$\Sigma = A\T A$ we can use this to compute an upper bound on $\beta_k$ via Lagrangian Relaxation as
 \begin{align*}
  \beta_k^2 & = \max \big\{ x\T A\T A x \suchthat \Norm{x}{2} \leq 1,\; \Norm{x}{0} \leq k \big\} \\
	    & \leq \inf_{\rho \geq 0} \max \big\{ x\T A\T A x - \rho 
            (\Norm{x}{0} - k) \suchthat \Norm{x}{2} \leq 1 \big\} \\
	    & = \inf_{\rho \geq 0} \phi(\rho) + \rho\, k. 
 \end{align*}

As one way to solve \eqref{phi} the semidefinite lifting technique is used again to compute an upper bound on $\phi$ via
\begin{equation}\label{Asp08Primal}\tag{A2-Primal}
 \begin{aligned}
  \max \quad & \sum_{i=1}^n \Tr(P_i \, B_i) \\
  \text{s.t.} \quad & \Tr(X) = 1, \\
  & X \succeq P_i \succeq 0 \quad \text{for } i = 1, \dots, n,
 \end{aligned}
\end{equation}
for $B_i = b_i\,b_i\T - \rho\, I$, where $b_i$ is the $i$-th column of the
square-root of $\Sigma$. For details, we refer the reader to~\cite{Asp08}. The corresponding dual problem can then be written as 
\begin{equation}\label{Asp08Dual}\tag{A2-Dual}
 \begin{aligned}
  \min \quad & \lambda_{\max}\Big(\sum_{i=1}^n Y_i \Big) \\
  \text{s.t.} \quad & Y_i \succeq B_i \quad \text{for } i = 1, \dots, n, \\
  & Y_i \succeq 0 \quad \ \ \text{for } i = 1, \dots, n,
 \end{aligned}
\end{equation}
which has only half as many variables as \eqref{Asp08Primal} and one fewer linear and semidefinite constraint. Then \eqref{Asp08Dual} can be solved for non-negative $\rho$, which should be smaller than $\Sigma_{11}$, because otherwise the
optimal solution for \eqref{phi} will always be $x=0$.

As shown in \cite{Asp08}, a lower bound for $\alpha_k$ can be computed via
\begin{equation}\tag{A2}\label{Asp08}
 \alpha_k^2 \geq \sup_{\rho \geq 0}\; \psi(\rho) - \rho\, k, 
\end{equation}
with 
\begin{equation*}
  \psi(\rho) = \min \big\{ x\T \Sigma\, x + \rho \Norm{x}{0} \suchthat \Norm{x}{2} \leq 1\big\}.
\end{equation*}
In this case problem \eqref{Asp08Primal} or equivalently \eqref{Asp08Dual} has to
be solved for $\Sigma = M I - A\T A$, where $M \in \R$ has to be big enough
to make $\Sigma$ positive semidefinite again. If $\Psi$ is the
corresponding optimal objective, this implies a lower bound on $\alpha_k$
in
\begin{equation*}
  \alpha_k^2 \geq \sup_{\rho \geq 0}\; (M - \Psi) - \rho\, k.
\end{equation*}
In \cite{AspBG14} explicit bounds on the gap between \eqref{Asp08Primal} and the penalized problem \eqref{phi} were derived using randomization arguments.

>>>>

\section{Contribution of this Work}

@@

The idea is that the optimal constant $\alpha_k^2$ in \eqref{lowerRIP} can be computed
by the non-convex quadratic optimization problem
\begin{equation}\label{QP}\tag{QP}
 \begin{aligned}
  \min \quad & \Norm{Ax}{2}^2  \\
  \text{s.t.} \quad & \Norm{x}{2}^2 = 1,  \\
  & \Norm{x}{0} \leq k. 
 \end{aligned}
\end{equation}
The corresponding maximization problem allows to compute $\beta_k^2$. D'Aspremont et al. \cite{Asp07} tackled the non-convex quadratic equality constraint 
$\Norm{x}{2}^2 = 1$ by the technique of semidefinite lifting, first introduced by Goemans and Williamson~\cite{GW95} 
for the max-cut problem. The idea is to use new matrix variables $X=xx\T$ for the quadratic terms, where the condition $X=xx\T$ is enforced by the equivalent constraints $X \succeq 0$ and $\Rk(X) = 1$, where, as usual, 
$X \succeq 0$ means that $X$ is symmetric (shortly written $X \in S_n$) and positive semidefinite. The objective and the normalization-constraint can then be expressed as  $\Norm{Ax}{2}^2 = x\T A\T Ax =
\Tr(A\T A X)$ and $\Norm{x}{2}^2 = \sum_{i=1}^n x_i^2 = \Tr(X) = 1$.

In the next step, the inequality $\Norm{x}{0} \leq k$ is substituted by the weaker constraint $\ones\T
\abs{X} \ones \leq k$, which follows from the fact that $\ones\T\abs{X} \ones = \Norm{\vect(X)}{1}\leq \sqrt{\Norm{\vect(X)}{0}}\, \Norm{\vect(X)}{2}$ and $\Norm{\vect(X)}{0}=\Norm{x}{0}^2$ in the rank-1 case, where $\vect(X)$ is a vector
in $\R^{n^2}$ consisting of all entries of $X$. 
Relaxing the non-convex rank constraint finally leads to the following SDP:
\begin{equation}\label{Asp07}\tag{A1}
 \begin{aligned}
  \min \quad & \Tr(A\T A X) \\
  \text{s.t.} \quad & \Tr(X) = 1, \\
  & \ones\T \abs{X} \ones \leq k, \\
  & X \succeq 0. \nonumber
 \end{aligned}
\end{equation}

Another SDP relaxation was proposed by d'Aspremont et al. in \cite{Asp08}. Here,
\begin{equation}\label{phi}
  \phi(\rho) = \max \big\{ x\T \Sigma\, x - \rho \Norm{x}{0} \suchthat \Norm{x}{2} \leq 1 \big\},
\end{equation}
originally stemming from sparse principal component analysis, is approximately solved for a symmetric positive semidefinite matrix $\Sigma$. For
$\Sigma = A\T A$ we can use this to compute an upper bound on $\beta_k$ via Lagrangian Relaxation as
 \begin{align*}
  \beta_k^2 & = \max \big\{ x\T A\T A x \suchthat \Norm{x}{2} \leq 1,\; \Norm{x}{0} \leq k \big\} \\
	    & \leq \inf_{\rho \geq 0} \max \big\{ x\T A\T A x - \rho 
            (\Norm{x}{0} - k) \suchthat \Norm{x}{2} \leq 1 \big\} \\
	    & = \inf_{\rho \geq 0} \phi(\rho) + \rho\, k. 
 \end{align*}

As one way to solve \eqref{phi}, the semidefinite lifting technique is used again to compute an upper bound on $\phi$ via
\begin{equation}\label{Asp08Primal}\tag{A2-Primal}
 \begin{aligned}
  \max \quad & \sum_{i=1}^n \Tr(P_i \, B_i) \\
  \text{s.t.} \quad & \Tr(X) = 1, \\
  & X \succeq P_i \succeq 0 \quad \text{for } i = 1, \dots, n,
 \end{aligned}
\end{equation}
for $B_i = b_i\,b_i\T - \rho\, I$, where $b_i$ is the $i$-th column of the
square-root of $\Sigma$. For details, we refer the reader to~\cite{Asp08}. The corresponding dual problem can then be written as 
\begin{equation}\label{Asp08Dual}\tag{A2-Dual}
 \begin{aligned}
  \min \quad & \lambda_{\max}\Big(\sum_{i=1}^n Y_i \Big) \\
  \text{s.t.} \quad & Y_i \succeq B_i \quad \text{for } i = 1, \dots, n, \\
  & Y_i \succeq 0 \quad \ \ \text{for } i = 1, \dots, n,
 \end{aligned}
\end{equation}
which has only half as many variables as \eqref{Asp08Primal} and one fewer linear and semidefinite constraint. Then \eqref{Asp08Dual} can be solved for non-negative $\rho$, which should be smaller than $\Sigma_{11}$, because otherwise the
optimal solution for \eqref{phi} will always be $x=0$.

As shown in \cite{Asp08}, a lower bound for $\alpha_k$ can be computed via
\begin{equation}\tag{A2}\label{Asp08}
 \alpha_k^2 \geq \sup_{\rho \geq 0}\; \psi(\rho) - \rho\, k, 
\end{equation}
with 
\begin{equation*}
  \psi(\rho) = \min \big\{ x\T \Sigma\, x + \rho \Norm{x}{0} \suchthat \Norm{x}{2} \leq 1\big\}.
\end{equation*}
In this case problem \eqref{Asp08Primal} or equivalently \eqref{Asp08Dual} has to
be solved for $\Sigma = M I - A\T A$, where $M \in \R$ has to be big enough
to make $\Sigma$ positive semidefinite. If $\Psi$ is the
corresponding optimal objective, this implies the following lower bound:
\begin{equation*}
  \alpha_k^2 \geq \sup_{\rho \geq 0}\; (M - \Psi) - \rho\, k.
\end{equation*}
In \cite{AspBG14} explicit bounds on the gap between \eqref{Asp08Primal} and the penalized problem \eqref{phi} were derived using randomization arguments.
